---
title: "RF sumirano"
output: html_document
---
```{r}
library("Biostrings")
library("Biobase")
library("dplyr")
library("plyr")
library("stringr")
library("coRdon")
library("randomForest")
library("Boruta")
library("RRF")
library("ipred")
library("gage")
library("ggplot2")
library("corrplot")
library("data.table")
library("pheatmap")
```



```{r}
samples <- readRDS("C:\\Users\\Maja\\Desktop\\Eva - diplomski\\samples\\all_samples.rds")
```


Svi uzorci kao trening set:
```{r}
sampleSet <- samples[samples$gt_1 > 0, c("all", "gt_1", "enrich", "A", "samples", "category")]
sampleSet <- melt(sampleSet, id.var = c("samples", "category"))
sampleSet$colnames <- paste(sampleSet$category, sampleSet$variable, sep = "_")
sampleSet <- sampleSet[, c(1, 4, 5)]

sample_names <- unique(sampleSet$samples)
condition <- factor(substr(sample_names, 1, 1))

sampleSet <- dcast(sampleSet, samples ~ colnames, fill = 0)
sampleSet$condition <- condition
rownames(sampleSet) <- sample_names
sampleSet <- sampleSet[, -1]

#sampleSet <- sampleSet[, order(apply(sampleSet, 2, sd), decreasing = TRUE)[1:5000]]
```

```{r}
random_forest <- RRF(condition ~ ., data = sampleSet, ntree = 1500, importance = TRUE, localImp = TRUE, proximity = TRUE, replace = TRUE, mtry = 1500)

pheatmap(random_forest$proximity, cellheight = 4, cellwidth = 4, fontsize = 4)
pheatmap(random_forest$proximity, cellheight = 6, cellwidth = 6, fontsize = 6)
#pheatmap(random_forest$proximity, cellheight = 10, cellwidth = 10)

plot(random_forest)
```





```{r}
rf10 <- RRF(condition ~ ., data = sampleSet, ntree = 10000, importance = TRUE, localImp = TRUE, proximity = TRUE, replace = TRUE, mtry = 2500)
random_forest <- RRF(condition ~ ., data = sampleSet, ntree = 1500, importance = TRUE, localImp = TRUE, proximity = TRUE, replace = TRUE, mtry = 2500)

#pheatmap(random_forest$proximity, cellheight = 4, cellwidth = 4, fontsize = 4)
#pheatmap(random_forest$proximity, cellheight = 6, cellwidth = 6, fontsize = 6)
#pheatmap(random_forest$proximity, cellheight = 10, cellwidth = 10)

plot(random_forest)

pheatmap(random_forest$proximity, cellheight = 4, cellwidth = 4, fontsize = 4)

```








MSE plot:
```{}
error_rate <- function(n){
  return(plot(random_forest)[n])
}

error <- data.table(ntrees = 1:2000)
error[, ER := error_rate(ntrees), by = ntrees]

error
```




Importance:
  -prouci sta je tocno taj importance, kako se mjeri i moze li se sta zanimljivo iz ovog izvuci
```{r}
#random_forest$importance[which((random_forest$importance != 0) == TRUE, arr.ind = T), ]
rnames <- rownames(which((random_forest$importance != 0) == TRUE, arr.ind = T))
rnames <- unlist(str_split(rnames, "_"))
importantKOs <- unique(grep("^K\\d+$", rnames, value = TRUE))
```

```{r}
library("KEGGREST")
#which(unique(importantKOs) %in% c("K09687"))
KOs <- data.table(KO = importantKOs[-62])
KOs[, name := (keggGet(KO)[[1]])$NAME, by = 1:nrow(KOs)]
KOs[, definition := (keggGet(KO)[[1]])$DEFINITION, by = 1:nrow(KOs)]

#saveRDS(KOs, "C:\\Users\\Maja\\Desktop\\Eva - diplomski\\samples\\KO_anotations.rds")
```






Metabolic Module Identification

Preparation of reference pathways. We will be using the KO-annotated metabolic pathways.
```{r}
path.set <- kegg.gsets("ko")
ko.gs <- path.set$kg.sets
```

Extraction of fold-change (M) values for the samples.
```{r}
gage_dt <- samples[, c("category", "M", "samples")]
gage_dt <- melt(gage_dt, id.var = c("samples", "category"))
gage_dt <- gage_dt[, -3]
gage_dt <- dcast(gage_dt, category ~ samples, fill = 0)
gage_dt <- as.data.frame(gage_dt)
rownames(gage_dt) <- gage_dt$category
gage_dt <- gage_dt[, -1]
```

GAGE run
```{r}
pathwayEnrich <- gage(gage_dt, gsets = ko.gs, ref = 1:80, compare = "unpaired")
kegg.sig <- sigGeneSet(pathwayEnrich)
```


Which genes are up- or downregulated in LD samples compred to HD samples:
-upregulated:
```{r}
upregulated <- data.frame(kegg.sig$greater)
upregulated_pathways <- rownames(upregulated)
```

-downregulated:
```{r}
downregulated <- data.frame(kegg.sig$less)
downregulated_pathways <- rownames(downregulated)
```









Clusters:
```{r}
cl1 <- c("LD49", "LD41", "LD48")
cl1 <- which(colnames(gage_dt) %in% cl1)

cl2 <- c("HD74", "HD9", "HD56", "HD50")
cl2 <- which(colnames(gage_dt) %in% cl2)

cl3 <- c("HD25", "HD14", "HD42", "HD46", "HD64")
cl3 <- which(colnames(gage_dt) %in% cl3)

cl4 <- c("HD16", "HD22", "HD39", "HD23", "HD65")
cl4 <- which(colnames(gage_dt) %in% cl4)

cl5 <- c("LD49", "LD41", "LD48", "LD12", "LD7", "LD29", "LD27", "LD19", "LD13", "LD18", "LD15", "LD42", "HD3", "HD47")
cl5 <- which(colnames(gage_dt) %in% cl5)

  
```


GAGE run
```{r}
pathwayEnrich <- gage(gage_dt, gsets = ko.gs, ref = cl5, compare = "unpaired")
kegg.sig <- sigGeneSet(pathwayEnrich)
```


Which genes are up- or downregulated in LD samples compred to HD samples:
-upregulated:
```{r}
upregulated <- data.frame(kegg.sig$greater)
upregulated_pathways <- rownames(upregulated)
upregulated_pathways
```

-downregulated:
```{r}
downregulated <- data.frame(kegg.sig$less)
downregulated_pathways <- rownames(downregulated)
downregulated_pathways
```







PCA plots:
```{r}
library(ggbiplot)
PCA <- prcomp(random_forest$proximity, scale. = TRUE)

PCAplot <- ggbiplot(PCA, ellipse = TRUE, var.axes = FALSE, groups = factor(substr(rownames(random_forest$proximity), 1, 1))) + theme_classic()
PCAplot

clusters <- kmeans(random_forest$proximity, 2)
cl1 <- names(which(clusters$cluster == 1))
cl2 <- names(which(clusters$cluster == 2))
cl1 <- which(colnames(gage_dt) %in% cl1)
```


Metabolic Module Identification

Preparation of reference pathways. We will be using the KO-annotated metabolic pathways.
```{r}
path.set <- kegg.gsets("ko")
ko.gs <- path.set$kg.sets
```

Extraction of fold-change (M) values for the samples.
```{r}
gage_dt <- samples[, c("category", "M", "samples")]
gage_dt <- melt(gage_dt, id.var = c("samples", "category"))
gage_dt <- gage_dt[, -3]
gage_dt <- dcast(gage_dt, category ~ samples, fill = 0)
gage_dt <- as.data.frame(gage_dt)
rownames(gage_dt) <- gage_dt$category
gage_dt <- gage_dt[, -1]
```

GAGE run
```{r}
pathwayEnrich <- gage(gage_dt, gsets = ko.gs, ref = cl1, compare = "unpaired")
kegg.sig <- sigGeneSet(pathwayEnrich)
```


Which genes are up- or downregulated in LD samples compred to HD samples:
-upregulated:
```{r}
upregulated <- data.frame(kegg.sig$greater)
upregulated_pathways <- rownames(upregulated)
upregulated_pathways
```

-downregulated:
```{r}
downregulated <- data.frame(kegg.sig$less)
downregulated_pathways <- rownames(downregulated)
downregulated_pathways
```









PCA plots:
```{r}
library(ggbiplot)
PCA <- prcomp(random_forest$proximity, scale. = TRUE)

PCAplot <- ggbiplot(PCA, ellipse = TRUE, var.axes = FALSE, groups = factor(substr(rownames(random_forest$proximity), 1, 1))) + theme_classic()
PCAplot

clusters <- kmeans(random_forest$proximity, 4)
cl1 <- names(which(clusters$cluster == 1))
cl1 <- which(colnames(gage_dt) %in% cl1)

cl2 <- names(which(clusters$cluster == 2))
cl2 <- which(colnames(gage_dt) %in% cl2)

cl3 <- names(which(clusters$cluster == 3))
cl3 <- which(colnames(gage_dt) %in% cl3)

cl4 <- names(which(clusters$cluster == 4))
cl4 <- which(colnames(gage_dt) %in% cl4)

```


Metabolic Module Identification

Preparation of reference pathways. We will be using the KO-annotated metabolic pathways.
```{r}
path.set <- kegg.gsets("ko")
ko.gs <- path.set$kg.sets
```

Extraction of fold-change (M) values for the samples.
```{r}
gage_dt <- samples[, c("category", "M", "samples")]
gage_dt <- melt(gage_dt, id.var = c("samples", "category"))
gage_dt <- gage_dt[, -3]
gage_dt <- dcast(gage_dt, category ~ samples, fill = 0)
gage_dt <- as.data.frame(gage_dt)
rownames(gage_dt) <- gage_dt$category
gage_dt <- gage_dt[, -1]
```

GAGE run
```{r}
pathwayEnrich <- gage(gage_dt, gsets = ko.gs, ref = cl4, compare = "unpaired")
kegg.sig <- sigGeneSet(pathwayEnrich)
```


Which genes are up- or downregulated in LD samples compred to HD samples:
-upregulated:
```{r}
upregulated <- data.frame(kegg.sig$greater)
upregulated_pathways <- rownames(upregulated)
upregulated_pathways
```

-downregulated:
```{r}
downregulated <- data.frame(kegg.sig$less)
downregulated_pathways <- rownames(downregulated)
downregulated_pathways
```
















