---
title: "nj"
output: html_document
---

```{r}
library("Biostrings")
library("Biobase")
library("dplyr")
library("plyr")
library("stringr")
library("coRdon")
library("randomForest")
library("Boruta")
library("RRF")
library("ipred")
library("gage")
library("ggplot2")
library("corrplot")
library("data.table")
```


######

Priprema uzoraka:
```{}
my_fasta_file_folder <- "C:\\Users\\Maja\\Desktop\\Eva - diplomski\\samples\\60uzoraka"
uzorci <- readSet(my_fasta_file_folder, prepend.filenames = T)
uzorci <- codonTable(uzorci)
```

```{}
condition <- vector()
sample_names <- vector()
ct_set <- list()
for (name in unique(str_extract(getID(uzorci), "[HL]D\\d+"))) {
  sample <- subset(uzorci, str_extract(getID(uzorci), "[HL]D\\d+") == name)
  melp <- MELP(sample, ribosomal = TRUE)
  ct <- crossTab(getKO(sample), as.numeric(melp), threshold = 1L)
  condition <- c(condition, rep(substr(getID(sample[1]), 1, 1), nrow(contable(ct))))
  sample_names <- c(sample_names, rep(str_extract(getID(sample)[1], "[HL]D\\d+"), nrow(contable(ct))))
  ct_set <- append(ct_set, ct)
}
```

```{}
all_enr <- lapply(ct_set, enrichment)
all_enr_data <- lapply((all_enr), pData)
```

```{}
samples <- do.call("rbind", all_enr_data)
samples$samples <- sample_names
samples$condition <- condition
```

```{}
saveRDS(samples, "C:\\Users\\Maja\\Desktop\\Eva - diplomski\\samples\\60samples.rds")
```

#####




```{r}
samples <- readRDS("C:\\Users\\Maja\\Desktop\\Eva - diplomski\\samples\\60samples.rds")
```


Training set:
```{r}
#promijeni ovisno o broju uzoraka   
HDs <- unique(samples$samples)[which(substr(unique(samples$samples), 1, 1) == "H")]
HDtrain_samples <- HDs[sample(1:30, 30 * 0.8)]
HDtrain <- samples[samples$samples %in% HDtrain_samples, ]

LDs <- unique(samples$samples)[which(substr(unique(samples$samples), 1, 1) == "L")]
LDtrain_samples <- LDs[sample(1:30, 30 * 0.8)]
LDtrain <- samples[samples$samples %in% LDtrain_samples, ]


train <- rbind(HDtrain, LDtrain)
```

Test set:
```{r}
HDtest_samples <- HDs[!(HDs %in% HDtrain_samples)]
HDtest <- samples[samples$samples %in% HDtest_samples, ]

LDtest_samples <- LDs[!(LDs %in% LDtrain_samples)]
LDtest <- samples[samples$samples %in% LDtest_samples, ]

test <- rbind(HDtest, LDtest)
```



Random Forest Classification
```{r}
#training set:
train_gt1 <- which(train$gt_1 > 0)

trainSet <- train[train_gt1, 2:8]
trainClasses <- factor(samples$condition[train_gt1])

#test set:
test_gt1 <- which(test$gt_1 > 0)

testSet <- test[test_gt1, 2:8]
testClasses <- factor(samples$condition[test_gt1])
```



Kada probam isfiltrirati:

Random Forest Classification
```{r}
#training set:
train_gt1 <- which(train$gt_1 > 0 & train$pvals < 0.05)

trainSet <- train[train_gt1, 2:8]
trainClasses <- factor(train$condition[train_gt1])

#test set:
test_gt1 <- which(test$gt_1 > 0 & test$pvals < 0.05)

testSet <- test[test_gt1, 2:8]
testClasses <- factor(test$condition[test_gt1])
```



```{r}
getImpRRF <- function(x, y, ...) {
rf <- RRF(x, y, importance = TRUE, keep.forest = FALSE, ntree = 20, flagReg = 1)  #1000
imp <- rf$importance[, 1]
imp/max(imp)
}
```

```{r}
bo <- Boruta(trainSet, trainClasses, maxRuns = 20, getImp = getImpRRF)   #200
relevantKOs <- names(bo$finalDecision[bo$finalDecision %in% c("Tentative", "Confirmed")])
```


```{r}
realTrainingset <- trainSet[, relevantKOs]
rrf <- RRF(trainClasses ~ ., data = realTrainingset, ntree = 10, importance = TRUE, localImp = TRUE,
proximity = TRUE, replace = FALSE) #ntree = 2000
```

```{r}
rrf.pred <- predict(rrf, testSet, proximity = TRUE)
err <- mean(rrf.pred$predicted!=testClasses)
err
```











#staro 













Prediction of Expression Levels
```{r}
uzorci <- readRDS("C:\\Users\\Maja\\Desktop\\Eva - diplomski\\samples\\smp\\60samples.rds")
```

```{r}
condition <- str_extract(colnames(uzorci)[2:ncol(uzorci)], "[HL]D")
all_samples <- uzorci[, 2:ncol(uzorci)]
```

Enrichment Analysis
For a single sample, we first create a contingency table summarising counts 
of genes annotated to each KO category among all the genes in sample, 
and among those predicted to be highly expressed. We do this using 
`crossTab()` function, giving it a character vector of genes' annotations, 
and a numeric vector of their respective MELP values, and specifying that 
a subset of highly expressed reference genes should contain those genes 
that have MELP value grater than 1 (this is default value of threshold). 
```{r}
ct.fun <- function(sample){
  ct <- crossTab(uzorci$genes, sample, threshold = 1L)
  return(ct)
}

all_ct <- as.data.table(t(data.table(apply(all_samples, 2, ct.fun))))
```

```{r}
all_enr <- lapply(unlist(all_ct), enrichment)   #tu jos dodaj pmethod (def prvo kao funkciju)
all_enr_data <- lapply((all_enr), pData)
```

```{r}
samples <- do.call("rbind", all_enr_data)
sample_names <- rep(colnames(all_samples), each = nrow(all_samples))
KO_names <- uzorci$genes
rownames(samples) <- paste(sample_names, KO_names, sep = "-")
```





Training set:
```{r}
#promijeni ovisno o broju uzoraka   
HDs <- unique(sample_names)[which(substr(unique(sample_names), 1, 1) == "H")]
HDtrain_samples <- HDs[sample(1:30, 30 * 0.8)]
HDtrain <- samples[str_extract(rownames(samples), "HD\\d+") %in% HDtrain_samples, ]

LDs <- unique(sample_names)[which(substr(unique(sample_names), 1, 1) == "L")]
LDtrain_samples <- LDs[sample(1:30, 30 * 0.8)]
LDtrain <- samples[str_extract(rownames(samples), "LD\\d+") %in% LDtrain_samples, ]


train <- rbind(HDtrain, LDtrain)
```

Test set:
```{r}
HDtest_samples <- HDs[!(HDs %in% HDtrain_samples)]
HDtest <- samples[str_extract(rownames(samples), "HD\\d+") %in% HDtest_samples, ]

LDtest_samples <- LDs[!(LDs %in% LDtrain_samples)]
LDtest <- samples[str_extract(rownames(samples), "LD\\d+") %in% LDtest_samples, ]

test <- rbind(HDtest, LDtest)
```





Random Forest Classification
```{r}
#training set:
train_gt1 <- which(train$gt_1 > 0)

trainSet <- train[train_gt1, 2:8]
trainClasses <- factor(substr(rownames(train[train_gt1, ]), 1, 1))

#test set:
test_gt1 <- which(test$gt_1 > 0)

testSet <- test[test_gt1, 2:8]
testClasses <- factor(substr(rownames(train[test_gt1, ]), 1, 1))
```



```{r}
getImpRRF <- function(x, y, ...) {
rf <- RRF(x, y, importance = TRUE, keep.forest = FALSE, ntree = 20, flagReg = 1)  #1000
imp <- rf$importance[, 1]
imp/max(imp)
}
```

```{r}
bo <- Boruta(trainSet, trainClasses, maxRuns = 20, getImp = getImpRRF)   #200
relevantKOs <- names(bo$finalDecision[bo$finalDecision %in% c("Tentative", "Confirmed")])
```


```{r}
realTrainingset <- trainSet[, relevantKOs]
rrf <- RRF(trainClasses ~ ., data = realTrainingset, ntree = 10, importance = TRUE, localImp = TRUE,
proximity = TRUE, replace = FALSE) #ntree = 2000
```

```{r}
rrf.pred <- predict(rrf, testSet, proximity = TRUE)
err <- mean(rrf.pred$predicted!=testClasses)
err
```







```{r}
which((all_train_enr[[3]]$gt_1 > 1) == TRUE)
nj <- all_train_enr[[1]][sample(1:nrow(all_train_enr[[1]]), 20) ,]
mj <- all_test_enr[[1]][sample(1:nrow(all_test_enr[[1]]), 20) ,]
#nj$condition <- c(rep("HD", nrow(nj)/2), rep("LD", nrow(nj)/2)) 
nj$condition <- sample(c("HD", "LD"), 20, replace = TRUE) 
mj$condition <- sample(c("HD", "LD"), 20, replace = TRUE) 
#mj$condition <- c(rep("HD", nrow(nj)/2), rep("LD", nrow(nj)/2)) 
```



Random Forest Classification
```{r}
#training set:
train_gt1 <- which(nj$gt_1 > 0)

trainSet <- nj[train_gt1, 2:8]
trainClasses <- factor(nj$condition[train_gt1])

#test set:
test_gt1 <- which(mj$gt_1 > 0)

testSet <- mj[test_gt1, 2:8]
testClasses <- factor(mj$condition[test_gt1])
```


```{r}
getImpRRF <- function(x, y, ...) {
rf <- RRF(x, y, importance = TRUE, keep.forest = FALSE, ntree = 10, flagReg = 1)  #1000
imp <- rf$importance[, 1]
imp/max(imp)
}
```

```{r}
bo <- Boruta(trainSet, trainClasses, maxRuns = 20, getImp = getImpRRF)   #200
relevantKOs <- names(bo$finalDecision[bo$finalDecision %in% c("Tentative", "Confirmed")])
```


```{r}
realTrainingset <- trainSet[, relevantKOs]
rrf <- RRF(trainClasses ~ ., data = realTrainingset, ntree = 500, importance = TRUE, localImp = TRUE,
proximity = TRUE, replace = FALSE) #ntree = 2000
```

```{r}
rrf.pred <- predict(rrf, testSet, proximity = TRUE)
err <- mean(rrf.pred$predicted!=testClasses)
err
```





